Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2014-11-13 paper/main.tex
minor

A recent project~\cite{veristack14} has implemented and verified a
quantitative logic to reason about stack-space usage, and modified the
verified CompCert C compiler to translate C level bound to x86 stack
bounds.  This quantitative logic is also based on the potential method
but has very rudimentary support for automation that cannot
automatically derive symbolic bounds and is not based on efficient LP
solving.  In contrast, our main
contribution is an automatic amortized analysis for C that can
derive parametric bounds for loops
and recursive functions fully automatically.
%
In this article, we also use a more general quantitative Hoare
logic that is parametric in the resource of interest.

Our work has been inspired by type-based amortized resource analysis
for functional programs~\cite{Jost03,HoffmannH10,HoffmannAH12}.  Here,
we present the first automatic amortized resource analysis for
C. None of the existing techniques can handle the example programs we
describe in this work.
%
The automatic analysis of realistic C programs is enabled by three
major improvements over previous work.  First, we solved the
long-standing open problem of extending automatic amortized resource
analysis to compute bounds for programs that loop on (possibly
negative) integers without decreasing one individual number in each
iteration.  Second, we extended the analysis system to deal with
non-linear control flow that is introduced by \code{break} and
\code{return} statements.  Third, for the first time, we have combined
an automatic amortized analysis with a system for interactively
deriving bounds. In particular, recent systems~\cite{HoffmannS13} that
deal with integers and arrays cannot derive bounds that depend on
values in mutable locations, possibly negative integers, or on
differences between integers.

@@

We have demonstrated that our approach improves the state-of-the-art
in resource bound analysis for C programs in three ways. First, our
technique is naturally compositional, tracks size changes of
variables, and can abstractly specify the resource cost of functions
(\pref{sec:AAA}). Second, it is easily combinable with established
qualitative verification to guide semi-automatic bound derivation
(\pref{sec:anno}). Third, we have shown that the local inference rules
of the derivation system automatically produce easily checkable
certificates for the derived bounds (\pref{sec:soundness}).
%
Our system is the first amortized resource analysis for C programs.
It addresses the longstanding open problem of extending automatic
amortized resource analysis to compute bounds for programs that loop
on signed integers and to deal with non-linear control flow\ifshort{.}{ that is
introduced by break and return statements.}

>>>>

Our work has been inspired by type-based amortized resource analysis
for functional programs~\cite{Jost03,HoffmannH10,HoffmannAH12}.  Here,
we present the first automatic amortized resource analysis for
C. None of the existing techniques can handle the example programs we
describe in this work.
%
The automatic analysis of realistic C programs is enabled by three
major improvements over previous work.  First, we extended the
analysis system to associate potential with not just individual
program variables but also multivariate intervals
and, more generally, auxiliary variables.  Second, we solved the
long-standing open problem of extending automatic amortized resource
analysis to compute bounds for programs that loop on (possibly
negative) integers without decreasing one individual number in each
iteration.  Third, for the first time, we have combined an automatic
amortized analysis with a system for interactively deriving bounds. In
particular, recent systems~\cite{HoffmannS13} that deal with integers
and arrays cannot derive bounds that depend on values in mutable
locations, possibly negative integers, or on differences between
integers.

A recent project~\cite{veristack14} has implemented and verified a
quantitative logic to reason about stack-space usage, and modified the
verified CompCert C compiler to translate C level bound to x86 stack
bounds.  This quantitative logic is also based on the potential method
but has very rudimentary support for automation. It is not based on
efficient LP solving so it cannot automatically derive symbolic
bounds.  In contrast, our main contribution is an automatic amortized
analysis for C that can derive parametric bounds for loops and
recursive functions fully automatically.
%
In this article, we also use a more general quantitative Hoare
logic that is parametric in the resource of interest.

@@

We have demonstrated that our approach improves the state-of-the-art
in resource bound analysis for C programs in three ways. First, our
technique is naturally compositional, tracks size changes of
variables, and can abstractly specify the resource cost of functions
(\pref{sec:AAA}). Second, it is easily combinable with established
qualitative verification to guide semi-automatic bound derivation
(\pref{sec:anno}). Third, we have shown that the local inference rules
of the derivation system automatically produce easily checkable
certificates for the derived bounds (\pref{sec:soundness}).
%
Our system is the first amortized resource analysis for C programs.
It addresses the long-standing open problem of extending automatic
amortized resource analysis to compute bounds for programs that loop
on signed integers and to deal with non-linear control flow\ifshort{.}{ that is
introduced by break and return statements.}
